Generative Adversarial Networks

### 2017-12-05 cs6735/project/report.tex
edits

\begin{abstract}
    Within the past decade, deep generative networks have often fallen short, especially when compared to deep discriminative models. However, generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.} \citep{2014arXiv1406.2661G}, have proved themselves as a promising research avenue for deep generative models. GANs present a new architecture which allows for the unsupervised training of a generative model through backpropogation. Due to the current research effort, state-of-the-art results are now being achieved on a regular basis \citep{2017arXiv171010196K, 2017arXiv171111585W}; however, GANs still encounter instability during training and do not yet poccess the ability to produce completely realistic images. Nevertheless, these networks offer encouraging results for unsupervised learning, semi-supervised, reinforcement learning and several image generation tasks. This research paper will examine the motivations and challenges associated with GANs, their training process, architecture variations and some applications that are currently being explored.
\end{abstract}

@@

The process of training a generative model is very similar to a discriminative model. Using large amounts of data collected from a specific domain, we train the generative model to generate data from that domain. As these models typically have fewer parameters than that the number of data samples, they are forced to internalize some representation of the data \citep{genmodelingopenai}. Unlike supervised training, there exists no single desired output. Therefore, the problem becomes defining a cost function which forces the generative model to produce data more like that of the target domain \citep{genmodelingopenai}.

@@

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of an probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models was the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \textit{generative machines}. These new models implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

@@

A maximum likelihood cost is another option for training GANs \citep{2017arXiv170100160G}. It can be shown that that this cost function minimizes the Kullback-Leibler divergence between $p_{data}$ and $p_{model}$ instead of the Jensen-Shannon divergence. This cost estimator is useful as it allows for comparison with other generative models; however, it behaves similarly to the minimax cost function as seen in figure \ref{fig:CostFunc}. The cost function for this version becomes:

@@

\subsection{Algorithm}
The following algorithm presents the initial training algorithm as introduced by Goodfellow \textit{et al.} in 2014 \citep{2014arXiv1406.2661G}. The first step in the algorithm is to optimize the discriminator $D$. Many researchers initially held the belief that it would be optimal to train $D$ for several iterations before training the generator; however, this is still a controversial topic among researchers \citep{2017arXiv170100160G}. It is Goodfellow's opinion that each model should be trained simultaneously \citep{2017arXiv170100160G}.

@@

\subsection{Feature Matching}
Feature matching involves the definition of a new cost function for GANs to reduce the probability that the generator will not overtrain on the current discriminator \citep{2016arXiv160603498S}. Instead of optimizing on the output of the discriminator, the new objective requires the generator to minimize the difference between the expected value of some intermediate layer. The new objective function can be described as $\Big|\Big|\mathbb{E}_{x\sim p_{data}(x)}\Big[f(x)\Big] - \mathbb{E}_{z\sim p_{z}(z)}\Big[f(G(z))\Big]\Big|\Big|_2^2$ where $f(x)$ is the output of the activation function of some layer of the discriminator. Experimental evidence gathered has demonstrated the usefulness of this technique, especially when the training process becomes instable \citep{2016arXiv160603498S}. Furthermore, feature matching is useful for semi-supervised GAN applications as will later be described in section \ref{sec:semisupervised}.

@@

\subsection{Label Smoothing}
Label smoothing is a technique that was first introduced in the 1980s which involves replacing the 0 and 1 targets with \textit{smoother} values such as 0.1 and 0.9 respectively \citep{2016arXiv160603498S}. This technique is used to improve the networks resistance to adversarial examples. Specifically, it is recommended to use only one-sided label smoothing by replacing the 0.9 label instead of 1 for the real data label. If one alters the label for fake data, problems may arise when $p_{data}(x)$ nears $0$ and $p_{model}(x)$ becomes large for some sample $x$ \citep{2016arXiv160603498S}.

@@

\subsection{Inference GANs} \label{sec:inferencegan}
The first GAN architectures lacked the ability to map an sample $x$ to its associated latent vector $z$. The ability to do this is often referred to as having an \textit{inference mechanism} \citep{2017arXiv171007035C}. In 2016, two new architectures were independently proposed which both introduced this mechanism by adding an inference network \citep{2016arXiv160509782D, 2016arXiv160600704D}. This new architecture involves two generative models, a decoder and encoder, which work together to fool the discriminator. Instead of only receiving a sample $x$, the discriminator is also trained on a latent space vector $z$ and has to decide if the pair is a real sample and its encoding or generated sample and its associated latent space vector. This architecture allows the encoder to learn the mapping of $x$ to $z$. Furthermore, the encoder generates a set of continuous variables. That is, if the vector $z$ is of length $n$, then $n$ means and $n$ variances will be generated. This allows for a probability distribution for each value of $z$. However, the reconstructed images still remain of poor quality \citep{2017arXiv171007035C}.

\subsection{Progressive GANs} \label{sec:proggans}
Progressive GANs do not offer an architecture variation but a new training methodology. Introduced in October 2017 by Karras \textit{et al.} \citep{2017arXiv171010196K}, these networks incrementally increase the number of convolutions within the generator and discriminator. A new convolution is added after the previous generator converges until a limit is reached. This permits the generator to progressively discover the image distribution. This technique also increases training speed and improves training stability \citep{2017arXiv171010196K}. Combined with new normalization techniques, progressive GANs allow for much higher resolution images to be generated \citep{2017arXiv171010196K}. The experimental results received by Karras offer promising new directions for GAN research and applications.

@@

As mentioned in section \ref{sec:inferencegan}, it is possible to train an encoder to map a sample $x$ to a latent vector $z$. This is an extremely useful feature for exploring the latent space \citep{2017arXiv171007035C}. For example, one may discover \textit{concept vectors} such as "wearing glasses" or "wearing a hat" and apply them to a vector associated with a human. An example of this process is illustrated in figure \ref{fig:conceptvector}.

\begin{figure}
\centering
\includegraphics[scale=0.40]{smile_concept_vector}
\caption{The addition of the "smile" concept vector (far right) to a vector associated with a woman (far left). The images in-between were created through an interpolation of the two vectors. Example retrieved from Dumoulin \textit{et al.} \citep{2016arXiv160600704D}.}
\label{fig:conceptvector}
\end{figure}

\section{Applications} \label{sec:applications}
Although GANs are a relatively new in the field of machine learning, many different applications have already been discovered. Several of the reoccurring themes are elaborated within this section.

\subsection{Image Generation} \label{sec:img2img}
Image generation is extremely useful as many tasks intrinsically require the ability to generate realistic samples \citep{2017arXiv170100160G}. Moreover, image generator often requires the ability to model multi-modal outputs (see figure \ref{fig:multimodal}) \citep{2017arXiv170100160G}. GANs are well suited for both of these tasks. Most research conducted with respect the GANs attempts to improve the quality of synthesized images \citep{2017arXiv171007035C}. As such, several useful and fascinating image applications have already been found.

@@

\subsection{Semi-Supervised Learning} \label{sec:semisupervised}
There exist several techniques for training a classifier using a GAN. These methods all take advantage of the unsupervised aspect of GANs to implement semi-supervised learning. The ability to use GAN for semi-supervised learning also offers a quantitative method to rank different GAN models.

One method involves the use of transfer learning. After the GAN training process has completed, the discriminator may be used within classification tasks as a feature extractor. A new machine learning model, such as a support vector machine, is trained on the extracted features of some layer $l$ of the discriminator. Another transfer learning approach makes use of inference GANs. A new machine learning model is trained on the last three hidden layers of the of the encoder. These methods have both received state-of-the-art classification results \citep{2017arXiv171007035C}.

Another general approach to semi-supervised learning involves the transformation of the discriminator into a classifier. Given that there are $k$ different labels for some classification task, the new classifier will have $k + 1$ classes where the $k + 1$ class is associated with a fake sample. When a real sample is given to the discriminator, it attempts to classify the sample as one of the $k + 1$ classes; however, when a generated or unlabeled sample is given, the probabilities of the k classes are summed together to represent the probability of the sample being real. In this situation, the classifier reverts back to its discriminative behavior. Typical supervised models need more than $50,000$ samples whereas state-of-the-art performance can be achieved on labeled datasets ranging from $20$ to $8,000$ samples using this technique \citep{2017arXiv170100160G}.

@@

\section{Conclusion}
GANs are generative models which allow us to parametrize complicated probability distribution functions. They can be used to generate, modify and manipulate images. Furthermore, they allow for the use novel techniques within the context of reinforcement learning and semi-supervised learning. Today, they are already being applied to many problems \citep{genmodelingopenai}; however, the majority of their applications have yet to be discovered. As long as the research effort continues, GANs promise exciting new research avenues within the field of machine learning.

>>>>

\begin{abstract}
    Within the past decade, deep generative networks have often fallen short, especially when compared to deep discriminative models. However, generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.} \citep{2014arXiv1406.2661G}, have proved themselves as a promising research avenue for deep generative models. GANs present a new architecture which allows for the unsupervised training of a generative model through backpropogation. Due to the current research effort, state-of-the-art results are now being achieved on a regular basis \citep{2017arXiv171010196K, 2017arXiv171111585W}; however, GANs still encounter instability during training and do not yet possess the ability to produce completely realistic images. Nevertheless, these networks offer encouraging results for unsupervised learning, semi-supervised learning, reinforcement learning and several image generation tasks. This research paper will examine the motivations and challenges associated with GANs, their training process, architecture variations and some applications that are currently being explored.
\end{abstract}

@@

The process of training a generative model is very similar to a discriminative model. Using large amounts of data collected from a specific domain, we train the generative model to generate data from that domain. As these models typically have fewer parameters than the number of data samples, they are forced to internalize some representation of the data \citep{genmodelingopenai}. Unlike supervised training, there exists no single desired output. Therefore, the problem becomes defining a cost function which forces the generative model to produce data more like that of the target domain \citep{genmodelingopenai}.

@@

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of a probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models was the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \textit{generative machines}. These new models implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

@@

A maximum likelihood cost is another option for training GANs \citep{2017arXiv170100160G}. It can be shown that this cost function minimizes the Kullback-Leibler divergence between $p_{data}$ and $p_{model}$ instead of the Jensen-Shannon divergence. This cost estimator is useful as it allows for comparison with other generative models; however, it behaves similarly to the minimax cost function as seen in figure \ref{fig:CostFunc}. The cost function for this version becomes:

@@

\subsection{Algorithm}
The following algorithm presents the initial training algorithm as introduced by Goodfellow \textit{et al.} in 2014 \citep{2014arXiv1406.2661G}. The first step in the algorithm is to optimize the discriminator $D$. Many researchers initially held the belief that it would be optimal to train $D$ for several iterations before training the generator; however, this is still a controversial topic amongst researchers \citep{2017arXiv170100160G}. It is Goodfellow's opinion that each model should be trained simultaneously \citep{2017arXiv170100160G}.

@@

\subsection{Feature Matching}
Feature matching involves the definition of a new cost function for GANs to reduce the probability that the generator will overtrain on the current discriminator \citep{2016arXiv160603498S}. Instead of optimizing on the output of the discriminator, the new objective requires the generator to minimize the difference between the expected value of some intermediate layer. The new objective function can be described as $\Big|\Big|\mathbb{E}_{x\sim p_{data}(x)}\Big[f(x)\Big] - \mathbb{E}_{z\sim p_{z}(z)}\Big[f(G(z))\Big]\Big|\Big|_2^2$ where $f(x)$ is the output of the activation function of some layer of the discriminator. Experimental evidence gathered has demonstrated the usefulness of this technique, especially when the training process becomes instable \citep{2016arXiv160603498S}. Furthermore, feature matching is useful for semi-supervised GAN applications as will later be described in section \ref{sec:semisupervised}.

@@

\subsection{Label Smoothing}
Label smoothing is a technique that was first introduced in the 1980s which involves replacing the 0 and 1 targets with \textit{smoother} values such as 0.1 and 0.9 respectively \citep{2016arXiv160603498S}. This technique is used to improve the networks resistance to adversarial examples. Specifically, it is recommended to use only one-sided label smoothing by replacing the 1 with 0.9 for the real data label. If one alters the label for fake data, problems may arise when $p_{data}(x)$ nears $0$ and $p_{model}(x)$ becomes large for some sample $x$ \citep{2016arXiv160603498S}.

@@

\subsection{Inference GANs} \label{sec:inferencegan}
The first GAN architectures lacked the ability to map a sample $x$ to its associated latent vector $z$. The ability to do this is often referred to as having an \textit{inference mechanism} \citep{2017arXiv171007035C}. In 2016, two new architectures were independently proposed which both introduced this mechanism by adding an inference network \citep{2016arXiv160509782D, 2016arXiv160600704D}. This new architecture involved two generative models, a decoder and encoder, which work together to fool the discriminator. Instead of only receiving a sample $x$, the discriminator is also trained on a latent space vector $z$ and has to decide if the pair is a real sample and its encoding or generated sample and its associated latent space vector. This architecture allows the encoder to learn the mapping of $x$ to $z$. Furthermore, the encoder generates a set of continuous variables. That is, if the vector $z$ is of length $n$, then $n$ means and $n$ variances will be generated. This allows for a probability distribution for each value of $z$. However, the reconstructed images still remain of poor quality \citep{2017arXiv171007035C}.

\subsection{Progressive GANs} \label{sec:proggans}
Progressive GANs do not offer an architecture variation but a new training methodology. Introduced in October 2017 by Karras \textit{et al.} \citep{2017arXiv171010196K}, these networks incrementally increase the number of convolutions within the generator and discriminator. A new convolution is added after the previous generator converges. This permits the generator to progressively discover the image distribution. This technique also increases training speed and improves training stability \citep{2017arXiv171010196K}. Combined with new normalization techniques, progressive GANs allow for much higher resolution images to be generated \citep{2017arXiv171010196K}. The experimental results received by Karras offer promising new directions for GAN research and applications.

@@

As mentioned in section \ref{sec:inferencegan}, it is possible to train an encoder to map a sample $x$ to a latent vector $z$. This is an extremely useful feature for exploring the latent space \citep{2017arXiv171007035C}. For example, one may discover \textit{concept vectors} such as ``wearing glasses'' or ``wearing a hat'' and apply them to a vector associated with a human. An example of this process is illustrated in figure \ref{fig:conceptvector}.

\begin{figure}
\centering
\includegraphics[scale=0.40]{smile_concept_vector}
\caption{The addition of the ``smile'' concept vector (far right) to a vector associated with a woman (far left). The images in between were created through an interpolation of the two vectors. Example retrieved from Dumoulin \textit{et al.} \citep{2016arXiv160600704D}.}
\label{fig:conceptvector}
\end{figure}

\section{Applications} \label{sec:applications}
Although GANs are a relatively new idea within the field of machine learning, many different applications have already been discovered. Several of the reoccurring themes are elaborated within this section.

\subsection{Image Generation} \label{sec:img2img}
Image generation is extremely useful as many tasks intrinsically require the ability to generate realistic images \citep{2017arXiv170100160G}. Moreover, image generation often requires the ability to model multi-modal outputs (see figure \ref{fig:multimodal}) \citep{2017arXiv170100160G}. GANs are well suited for both of these tasks. Most research conducted with respect the GANs attempts to improve the quality of synthesized images \citep{2017arXiv171007035C}. As such, several useful and fascinating image applications have already been found.

@@

\subsection{Semi-Supervised Learning} \label{sec:semisupervised}
There exist several techniques for training a classifier using a GAN. These methods all take advantage of the unsupervised aspect of GANs to implement semi-supervised learning. The ability to use GANs for semi-supervised learning also offers a quantitative method to rank different GAN models.

One method involves the use of transfer learning. After the GAN training process has completed, the discriminator may be used within classification tasks as a feature extractor. A new machine learning model, such as a support vector machine, may be trained on the extracted features of some layer $l$ of the discriminator. Another transfer learning approach makes use of inference GANs. A new machine learning model is trained on the last three hidden layers of the of the encoder. These methods have both received state-of-the-art classification results \citep{2017arXiv171007035C}.

Another general approach to semi-supervised learning involves the transformation of the discriminator into a classifier. Given that there are $k$ different labels for some classification task, a new classifier is created that will have $k + 1$ classes where the $k + 1$ class is associated with a fake sample. When a real sample is given to the discriminator, it attempts to classify the sample as one of the $k + 1$ classes; however, when a generated or unlabeled sample is given, the probabilities of the k classes are summed together to represent the probability of the sample being real. In this situation, the classifier reverts back to its discriminative behavior. Typical supervised models need more than $50,000$ samples whereas state-of-the-art performance can be achieved on labeled datasets ranging from $20$ to $8,000$ samples using this technique \citep{2017arXiv170100160G}.

@@

\section{Conclusion}
GANs are generative models which allow us to parametrize complicated probability distribution functions. They can be used to generate, modify and manipulate images. Furthermore, they allow for the use novel techniques within the context of reinforcement learning and semi-supervised learning. Today, they are already being applied to many problems \citep{genmodelingopenai}; however, the majority of their applications have yet to be discovered. GANs promise numerous exciting new research avenues within the field of machine learning.
